---
title: "sensRICLPM_example"
author: "Lina Kramer"
date: "4/26/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(magrittr)
library(purrr)
library(lavaan)
library(tidyr)
library(dplyr)


library(sensRICLPM)

set.seed(155)
```

# sensRICLPM sensitivity analysis example


## Simulate data according to Mulder (2022)

**3 waves of data (stationary & symmetric): defining population parameters:**

*Autoregressive and cross-lagged parameters:*

$\alpha = .4$, 
$\beta = .2$, 
$\gamma = .2$,
$\delta = .3$

*Within variances & covariances at wave 1:*

$VAR(WX) = 1$,  
$VAR(WY) = 1$,  
$COR(WX,WY) = .3$ 

*Unique variances & covariances at waves 2 & 3:*

The unique variances and covariances were determinded using Kim and Nelson's (1999, p.27) expression of the unconditional covariance matrix of a stationary process as a function of the lagged effects and the residual variance covariance matrix:
$vec(\Psi) =(I-\Phi\otimes \Phi)vec(\Sigma)$

    
where $$\Sigma = \begin{bmatrix}
  \sigma_{w_{X1}} & \sigma_{w_{X1},w_{Y1}} \\
  \sigma_{w_{X1},w_{Y1}} & \sigma_{w_{Y1}} 
    \end{bmatrix}$$, 
    
and $$\Phi =\begin{bmatrix}
    \alpha & \gamma \\
    \beta & \delta 
\end{bmatrix}$$


```{r}
# simulate data according to Mulder (2022)'s amotivation example

# unique covariance calculation according to Kim & Nelson:


Phi <- matrix(c(.4, .2, .2, .3), ncol=2, byrow=F) # Phi matrix alpha, beta, gamma, delta

wSigma <- matrix(c(1, .3, .3, 1), ncol=2, byrow=TRUE) # Matrix of (co-) variances of within components at wave 1  ;  var_wx1, cov_w1, cov_w1, var_wy1

# using Kim & Nelson's (1999, p. 27) formula:
Psi <- matrix((diag(length(wSigma)) - Phi %x% Phi) %*% c(wSigma), nrow = nrow(Phi))

Psi


```
```{r}
# compute measurement error variances needed 

# for a measurement error proportion of .3 in X
(.3*1 + .3*1)/(1 - .3) #  0.8571429

# for a measurement error proportion of .4 in Y
(.5*1+ .5*1)/(1 - .5) # 2.5
```


```{r}

# generate data 

populationsynt1 <- '
  
  ## Create between components (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 
  RIy =~ 1*y1 + 1*y2 + 1*y3
  
  ## Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  
  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3

  ##autogressive effects between the within-person centered variables.
  
  ##alpha == 0.4
  
  wx2 ~ 0.4*wx1
  wx3 ~ 0.4*wx2
  
  ##delta == 0.3
  
  wy2 ~ 0.3*wy1
  wy3 ~ 0.3*wy2
  
  ## cross-lagged effects
  
  ##gamma == 0.2
  
  wx2 ~ 0.2*wy1
  wx3 ~ 0.2*wy2
  
  ##beta == 0.2
  
  wy2 ~ 0.2*wx1
  wy3 ~ 0.2*wx2

  ## covariance between the within-person centered variables at the first wave
  wx1 ~~ 0.3*wy1 
  
  ## covariances between the residuals of the within-person centered variables (the innovations)
  
  ##covu == 0.112
  wx2 ~~ 0.112*wy2
  wx3 ~~ 0.112*wy3

  
  ## variances and covariance of the random intercepts. 
  
  ##varRI == 1
  ##covRI == .5
  RIx ~~ 1*RIx
  RIy ~~ 1*RIy
  RIx ~~ 0.5*RIy
  
  ## (residual) variance of the within-person centered variables.
  
  wx1 ~~ 1*wx1 # Variances in first wave
  wy1 ~~ 1*wy1 
  
  wx2 ~~  0.752*wx2 # Residual variances
  wy2 ~~ 0.834*wy2 
  wx3 ~~  0.752*wx3 
  wy3 ~~ 0.834*wy3 
    
  ## include ME variances
  
  ##xx == 0.857
  x1 ~~ 0.857*x1
  x2 ~~ 0.857*x2
  x3 ~~ 0.857*x3

  ##yy == 0.8571429
  y1 ~~ 0.857*y1
  y2 ~~ 0.857*y2
  y3 ~~ 0.857*y3
'

dataset1 <- simulateData(model=populationsynt1, meanstructure = F,
                      sample.nobs = 2000)
cov(dataset1) # check

```

```{r}
# run sensitivity analysis with sensRICLPM function

exampleresult1 <-sensRICLPM(dataset1, ME_prop_x = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), ME_prop_y = c(0, 0.1, 0.2, 0.3, 0.4, 0.5))

exampleresult1$p2 # get sensitivity plots

```


```{r}

# sensitivity analysis table (produces table 1)

restable <- exampleresult1$outputtable %>% filter(admissible== "TRUE"  & MEpropx == MEpropy & parameter %in% c("alpha", "beta", "gamma", "delta", "RIX", "RIY", "RIcov", "wX", "wY", "wcov")) %>% select(MEpropy, parameter, wave, est, cilower, ciupper) # se, pvalue, cilower, ciupper) 

restable <- restable %>% pivot_wider(names_from = MEpropy, values_from = c(est, ciupper, cilower))
restable$CIa <- paste0("[", round(restable$cilower_0, 3), ", ", round(restable$ciupper_0, 3), "]")
restable$CIb <- paste0("[", round(restable$cilower_0.1, 3), ", ", round(restable$ciupper_0.1, 3), "]")
restable$CIc <- paste0("[", round(restable$cilower_0.2, 3), ", ", round(restable$ciupper_0.2, 3), "]")
restable$CId <- paste0("[", round(restable$cilower_0.3, 3), ", ", round(restable$ciupper_0.3, 3), "]")

restable <- restable %>% select(parameter, wave, est_0, CIa,  est_0.1, CIb, est_0.2, CIc, est_0.3, CId)

restable$est_0 <- round(restable$est_0, 3)
restable$est_0.1 <- round(restable$est_0.1, 3)
restable$est_0.2 <- round(restable$est_0.2, 3)
restable$est_0.3 <- round(restable$est_0.3, 3)

#restable <- restable %>% knitr::kable(booktabs = T, format = "latex")

restable
```
```{r}
saveRDS(restable, "table1.rds")
```


